---
title: "In class assignment week 2, part 2. A worked example using control flow (for loops, if statements, etc)"
author: "Ian Dworkin"
editor: "Garrett Nunn"
output: 
  html_document:
    keep_md: yes
    number_sections: yes
---

# Introduction
Let's do a little exercise integrating some of the things we have learned. Here are some Illumina HiSeq reads for one of our recent projects:

```{r}
read_1 <- "CGCGCAGTAGGGCACATGCCAGGTGTCCGCCACTTGGTGGGCACACAGCCGATGACGAACGGGCTCCTTGACTATAATCTGACCCGTTTGCGTTTGGGTGACCAGGGAGAACTGGTGCTCCTGC"

read_2 <- "AAAAAGCCAACCGAGAAATCCGCCAAGCCTGGCGACAAGAAGCCAGAGCAGAAGAAGACTGCTGCGGCTCCCGCTGCCGGCAAGAAGGAGGCTGCTCCCTCGGCTGCCAAGCCAGCTGCCGCTG"

read_3  <- "CAGCACGGACTGGGGCTTCTTGCCGGCGAGGACCTTCTTCTTGGCATCCTTGCTCTTGGCCTTGGCGGCCGCGGTCGTCTTTACGGCCGCGGGCTTCTTGGCAGCAGCACCGGCGGTCGCTGGC"
```

Question 1. what species are these sequences from?
1 - Drosophila melanogastor
2 - Dm as well 
3 - Same

Question 2. Put all three of these reads into a single object (a vector).  What class will the vector `reads` be? Check to make sure! How many characters are in each read (and why does `length()` not give you what you want.. try...)
```{r}
reads = c(read_1, read_2, read_3)
length(reads)
nchar(reads)
```
Because length is giving you the length of the list

Question 3. Say we wanted to print each character (not the full string) from read_1, how do we do this using a for loop? You may wish to look at a function like `strsplit()` to accomplish this (there are other ways.)

Replace the blanks.
```{r}
read_1_split <- strsplit(read_1, split = "", fixed = T)
splitReads =strsplit(reads, split = "", fixed = T)
for (char in read_1_split[[1]]){
  print(char)
}
```

Question 4. What kind of object does this return? How might we make it a character vector again?
  It returns a character object of length 1.
```{r}
rebuilt = as.character(read_1_split[[1]])
```
Question 5. How about if we wanted the number of occurrences of each base? Or better yet, their frequencies? You could write a loop, but I suggest looking at the help for the `table()` function... Also keep in mind that for for most objects `length()` tells you how many elements there are in a vector. For lists use `lengths()` (so you can either do this on a character vector or a list, your choice)
```{r}
freq = table(rebuilt)
freq[] = freq[]/length(rebuilt)

```

Question 6. How would you make this into a nice looking function that can work on either  a list or vectors of characters? (Still just for a single read)
```{r}
countFreq = function(input){
  freq <- as.character(strsplit(input, split = "", fixed = T)[[1]])
  freq = table(freq)/length(freq)
  return(freq)
}
countFreq(read_1)
```

Question 7. Now how can you modify your approach to do it for an arbitrary numbers of reads? You could use a loop or use one of the apply like functions (which one)?
```{r}
countFreqs = function(input){
  freqs=list()
  for(i in 1:length(input)){
    freq <- as.character(strsplit(input[i], split = "", fixed = T)[[1]])
    freq = table(freq)/length(freq)
    freqs= cbind(freqs,freq)
    colnames(freqs)[i]=i
  }
  return(freqs)
}
countFreqs(reads)
```
Question 8. Can you revise your function so that it can handle the input of either a string as a single multicharacter vector, **or** a vector of individual characters **or** a list? Try it out with the vector of three sequence reads (`reads`).

You Finished this in class for us. Basically just sapply my countFreq function

