---
title: "Visualizing Data"
author: "Garrett Nunn"
date: "November 18, 2018"
output:
  html_document:
    keep_md: yes
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(MASS)
library(lattice)
```
#Data Visualization
Exploratory Visualizations are for data analysis so you want to be able to see as much information as possible in a meaningful way.

Explantory visualizations are for when you haev already found the relavent data and would like to show it to someone else.
* Base graphics is the basic graphics plotting software in R
* Grid graphics is a more complicated version of this
* ggplot2 is for scrubs but is based off the grid graphics system

Base Graphics:
```{r}
plot(xVar, yVar,xlab = "label of x axis", ylab = "y axis label", main= "Title")
title("Title of the Plot")
```
Grid:
* Basically used by people who make r packages
Lattice:
* Makes it very easy to make a conditional scatterplot
```{r}
xyplot(yVar ~ xVar |conditionalVar, data = data)
```
*Other plots:*
There are also different kind of plots in Base such as:
  *'sunflowerplot()' shows an additional petal for every value at the same point in the graph. Useful when data points are repeated
  *'boxplot()' useful for seeing distribution over a given y value and either a few x values or categories
  *'mosaicplot()' shows a mosaice plot
*Base Plot can also be enhanced*
  * 'points()' adds points to the plot
  * 'lines()' adds lines that are usually curved
  * 'text()' adds labels
  * 'par()' allows you to set-up multiple plot arrays
A good general rule of thumb is pie charts are bad

##Characterizing a single variable
Numerical data should general conform to a gaussian distribution. The spread is the stdev.
* Histogram is just a grouping of number of things in each bin
* Density estimate should look like a bell curve roughly. Basically a smoothed histogram
* Normal QQ-plot sorted points should be close to the solid line if the data is gaussian distribution

##Histograms
* 'hist()' is a function that puts numerical values into bins and shows you the number frequency of them
* 'truehist()' is similar to 'hist()' but instead shows each bins approximate estimate

##Characterizing the relationship between two variables
* Scatterplots do this 'plot(yVar~xVar, data = mydata)'
* Sunflower plots do this as well
* When few x values, use a boxplot
* Mosaicplot when you have a ton of repeated data points
*More complex relationships*
* Bagplots shows a 2D version of distribution for two variables
* Correlation coefficient is the relation between numerical varibales and is between -1 and +1
  + corrplot() shows this relation across the entire dataset
  + Useful prelimary view
* Decision tree models
  + Shows the structure of the relation between the other variables
```{r}
theCorrelationMatrix = corr(myData)
corrplot(corr=theCorrelationMatrix, method = "ellipse")
```
Tree modeling:
```{r}
tree_model = rpart(var2Predict ~ ., data = myData)
```
* Careful with this var2predict is the variable you are trying to pre

##Graphics Parameters
* Par is a global graphics parameter so 'par(pch = 16)' would change all future graphs to a closed circle from the default open one
* Changing it locally just affects the one plot you are making

